Data Specialization for a Staged Lambda Calculus

### 2015-02-27 paper/implementation.tex
more small cleanups

We allow staging annotations around declarations. For example,
\begin{lstlisting}
1`atsigngrnd{
  datatype list = Empty | Cons of int * list
  fun part (...) = ...
}
fun qsStaged (...) = ...`
\end{lstlisting}
declares the \texttt{list} datatype and \texttt{part} function at world \bbonep,
by elaborating into:
\begin{lstlisting}
1`val grnd{Empty} = grnd{roll (inj ...)}
val grnd{Cons}  = grnd{fn (...) => roll (inj ...)}
val grnd{part}  = grnd{fn (...) => ...}
val qsStaged    = fn (...) => ...`
\end{lstlisting}

>>>>

We allow staging annotations around declarations. For example,
\begin{lstlisting}
3`atsigngrnd{`1`
  datatype list = Empty | Cons of int * list
  fun part (...) = ...
`3`}`1`
fun qsStaged (...) = ...`
\end{lstlisting}
declares the \texttt{list} datatype and \texttt{part} function at world \bbonep,
by elaborating into:
\begin{lstlisting}
1`val 3`grnd{`Empty3`}` = 3`grnd{`roll (inj ...)3`}`
val 3`grnd{`Cons3`}`  = 3`grnd{`fn (...) => roll (inj ...)3`}`
val 3`grnd{`part3`}`  = 3`grnd{`fn (...) => ...3`}`
val qsStaged    = fn (...) => ...`
\end{lstlisting}

### 2015-02-27 paper/semantics.tex
fixing various broken things
more small cleanups

\paragraph{First-stage evaluation at \bbonem\ ($\redonesym$)}
The difficulty of evaluating terms at \bbonem\ is that they contain both terms
at \bbonep\ (via $\curr$) and at \bbtwo\ (via $\fut$); the other evaluation
rules are ordinary call-by-value rules. When we encounter $\pure{e}$, we
evaluate $e$ as a monostaged term; \texttt{letg} and \texttt{caseg} simply reach
inside $\pure$ and evaluate like \texttt{let} and \texttt{case}.

@@

\paragraph{First-stage evaluation at \bbtwo\ ($\redtwosym$)}
Since terms at world \bbtwo\ represent second-stage computations, we should not
evaluate them yet. However, we do need to evaluate any first-stage computations
they contain (inside $\prev$), even if those computations are underneath
binders.

@@

A multistage function, such as \texttt{qsStaged} from \ref{sec:example}, has
type $A\to \fut (B \to C)$ at \bbonem.%
\footnote{We can rewrite \texttt{qsStaged} in this curried form, or apply a
higher-order currying function.}
To apply this function to arguments $\colmix a A$ and $\coltwo b B$, simply
evaluate the program:
\begin{lstlisting}
2`prev{`1`f a`2`} b`
\end{lstlisting}

>>>>

\paragraph{First-stage evaluation at \bbonem\ ($\redonesym$).}
The difficulty of evaluating terms at \bbonem\ is that they contain both terms
at \bbonep\ (via $\curr$) and at \bbtwo\ (via $\fut$); the other evaluation
rules are ordinary call-by-value rules. When we encounter $\pure{e}$, we
evaluate $e$ as a monostaged term; \texttt{letg} and \texttt{caseg} simply reach
inside $\pure$ and evaluate like \texttt{let} and \texttt{case}.

@@

\paragraph{First-stage evaluation at \bbtwo\ ($\redtwosym$).}
Since terms at world \bbtwo\ represent second-stage computations, we should not
evaluate them yet. However, we do need to evaluate any first-stage computations
they contain (inside $\prev$), even if those computations are underneath
binders.

@@

A multistage function, such as \texttt{qsStaged} from \ref{sec:examplea}, has
type $A\to \fut (B \to C)$ at \bbonem.%
\footnote{We can rewrite \texttt{qsStaged} in this curried form, or apply a
higher-order currying function.}
To apply this function to arguments $\colmix a A$ and $\coltwo b B$, simply
evaluate the program:
\begin{lstlisting}
2`prev{`1`f a`2`} b`
\end{lstlisting}
